---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

## Q1
## Loading and preprocessing the data
```{r, echo=TRUE}
setwd("C:/Users/Bhagyashree Deshmukh/Downloads")
activity<-read.csv("activity.csv")
```

```{r}
dim(activity)
names(activity)
head(activity)
str(activity)
summary(activity)
#total number of missing data
sum(is.na(activity$steps))
```

## Q2 and Q3
## What is mean total number of steps taken per day?
## mean and median of total number of ateps taken per day
## histogram of total number of steps taken each day

```{r}
library(ggplot2)
total.steps<-tapply(activity$steps, activity$date, FUN=sum, na.rm=TRUE)
mean(total.steps, na.rm=TRUE)
median(total.steps, na.rm=TRUE)
qplot(total.steps,geom="histogram",xlab="Total Steps",ylab="Counts",main="Total Steps Historgram")
png("plot1.1.png")
dev.off()
```

## Q4 and Q5
## What is the average daily activity pattern?
## The 5-minute interval that, on average, contains the maximum number of steps

```{r}
averages <- aggregate(x=list(steps=activity$steps), by=list(interval=activity$interval),FUN=mean, na.rm=TRUE)
ggplot(data=averages, aes(x=interval, y=steps)) +
    geom_line() +
    xlab("5-minute interval") +
    ylab("average number of steps taken")
```

```{r}
averages[which.max(averages$steps),]
```

## Q6
## Imputing missing values

```{r}
missing<- is.na(activity$steps)
table(missing)
```

# Replace each missing value with the mean value of its 5-minute interval
```{r}
fill.value <- function(steps, interval) {
    filled <- NA
    if (!is.na(steps))
        filled <- c(steps)
    else
        filled <- (averages[averages$interval==interval, "steps"])
    return(filled)
}
filled.activity <- activity
filled.activity$steps <- mapply(fill.value, filled.activity$steps, filled.activity$interval)
```

## Q7
## Now, using the filled data set, let's make a histogram of the total number of steps taken each day and calculate the mean and median total number of steps.
```{r}
total.steps <- tapply(filled.activity$steps, filled.activity$date, FUN=sum)
qplot(total.steps, binwidth=1000, xlab="total number of steps taken each day after replacing each missing value by mean")
mean(total.steps)
median(total.steps)
```

## Q8
## Are there differences in activity patterns between weekdays and weekends?

```{r}
weekday.or.weekend <- function(date) {
    day <- weekdays(date)
    if (day %in% c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday"))
        return("weekday")
    else if (day %in% c("Saturday", "Sunday"))
        return("weekend")
    else
        stop("invalid date")
}
filled.activity$date <- as.Date(filled.activity$date)
filled.activity$day <- sapply(filled.activity$date, FUN=weekday.or.weekend)
```

```{r}
averages <- aggregate(steps ~ interval + day, data=filled.activity, mean)
ggplot(averages, aes(interval, steps)) + geom_line() + facet_grid(day ~ .) +
    xlab("5-minute interval") + ylab("Number of steps")
```
